Enhancing \Hinf{} Norm Estimation using Local LPM/LRM Modeling: Applied to an AVIS\thanksref{thx}

### 2016-03-04 ch04-lrmhinf/paper.tex
Add some TODOs
Split out Avis photo
Improvements CH04

% \renewcommand{\glsgroupskip}{}
% \glsdisablehyper

\input{\thisDir/styles.inc.tex}

@@

To estimate $\gamma$, only $\Delta(\omega_{\star})$ is of interest, where
$ \omega_{\star} 
     \isdef 
       \arg
         \sup_{\omega} 
           \abs{\Delta(\omega)}
  \text{.}
  \label{eq:omega-star}
$
However, in a finite measurement time, non-parametric approaches only allow access to a finite and discrete grid of frequencies $\omega_k \in \DFTGrid$.
As such, \eqref{eq:HinfNormSISO} can be rewritten as
\begin{equation}
  \gamma \isdef \sup_{\omega} \abs{\Delta(\omega)}
         = \sup 
             \set{
                \ezbrace{\sup_{\omega \in \DFTGrid  } \abs{\Delta(\omega)}}{\gamma_{\mathrm{FRF}}}
                \;,\;
                \ezbrace{
                \sup_{\omega \in \InterGrid} \abs{\Delta(\omega)}}{\gamma_{\mathrm{IG}}}}
  \label{eq:gamma-split-grid}
\end{equation}
where $\Omega \isdef \OpenInterval{0}{\frac{2\pi}{\Ts}}$ is the measured frequency range for a sampling time $\Ts$.
Two cases can be distinguished:
\begin{enumerate}
  \item $\omega_{\star} \in \DFTGrid$ such that $\abs{\Delta(\omega_{\star})} = \gamma =\gamma_{\mathrm{FRF}}$ is observed directly on the discrete frequency grid,
  \item $\omega_{\star} \not\in \DFTGrid$, i.e. the extremal value $\abs{\Delta(\omega_{\star})}  = \gamma = \gamma_{\mathrm{IG}}$ is determined by the inter-grid behavior.
\end{enumerate}
The use of $\gamma_{\mathrm{FRF}}$ is standard in \gls{FRF}-based approaches, e.g.~\citep{vandeWal2002}.
Its expected value over different realizations of the input signal $\expectedValue{\gamma_{\mathrm{FRF}}} \leq \gamma$ due to the limited frequency resolution.
Such an underestimate may lead to overly optimistic controller designs.

@@

\section{\Hinf{} Norm Estimation Procedure}
\label{sec:LPMHinf}
To estimate $\gamma_{\mathrm{IG}}$, we propose to identify local parametric models in the frequency domain.
The frequency axis is partitioned into continuous segments $\Omega_k \isdef \OpenInterval{ \omega_{k-1} }{ \omega_{k+1} }$ in view of the local validity of the local models.
In particular, the inter-grid aspect of the $\Hinf$ norm is estimated as
\begin{equation}
  \gamma_{\mathrm{IG}} = 
    \sup \set{ \left. 
                \sup_{\omega \in \Omega_k} \abs{\LocalModel{\Delta}_k(\omega)}
                \right|
                k = 1,\ldots,N-1
                }
   \label{eq:gamma-inter-grid}
\end{equation}
where the $\LocalModel{\Delta}_{k}$ is the local model around the ($k$th) frequency $\omega_k$.
An example of such local models is shown in \figref{fig:interpol-quantities}.
Two alternative methods allow to evaluate~\eqref{eq:gamma-inter-grid} practically.

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.8\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \input{\thisDir/figs/interpolation-show.tikz}
  \caption{Illustration of the local models $\LocalModel{\Delta}_k$ around each frequency bin $\omega_k$ for $k\in\set{1,\ldots,6}$.
  The bottom part shows the equivalence of the frequency segments $\Omega_k$ and the intervals $\set{\omega \given{k_L(\omega) = k \vee k_R(\omega) = k}}$.}
\label{fig:interpol-quantities}
\end{figure}

@@

\section{Simulation}
\label{sec:simulation}
To allow the generalization of the results of the simulations, we first introduce a way to decompose complex systems into simpler systems.
The behavior of these simpler systems then allows to predict the behavior for these more complex systems.

Any real-rational system $\Delta$ can be decomposed in a parallel structure of lower-order sub-systems.
As a result, systems can often be well-represented, in a small frequency range, by a sum of a few low-order sub-systems (often first or second order for e.g. mechanical systems)~\citep[Section 2.2.2]{Gawronski2004}.
To estimate $\norm{\Delta}$ mainly the resonant sub-systems are of interest.
Assuming such resonances are sufficiently separated in the frequency,~\citep{Schoukens2013LPMerror,Geerardyn2013TIM,Gawronski2004} show that results for a single resonance are applicable to more complex systems.
This is done by considering only the dominant resonance -- \emph{viz.} the one that is the most relevant for the $\Hinf{}$ controller design.
By normalizing the conclusions for the time constant $\tau$ (or the $3\unit{dB}$ bandwidth), the results are generally valid for different values of the damping $\xi$.
Specifically, the system's $\tau \isdef (\xi\omega_{\mathrm{n}})^{-1}$ dictates the required measurement time~\citep{Schoukens2013LPMerror}, where $\xi$ is the relative damping and $\omega_{\mathrm{n}}$ the undamped natural frequency.

\subsection{A concrete simulation}
Consider a discrete time system $\Delta$ (see \figref{fig:exampleIntergrid}) as unmodeled dynamics in an output-error setting:
\begin{equation}
  \Delta(z) 
    =  \frac{0.45373 z + 0.44752}
            {z^2 - 1.0595 z + 0.96079}
  \label{eq:O2sysDT}
  \text{,}
\end{equation}
which has $\tau = 50 \unit{samples}$ (i.e. $\xi=0.02$ and $\omega_{\mathrm{n}} = 1 \unit{rad/S}$).
The input signal $u_{\Delta}$ is white Gaussian noise with unit variance and the disturbing noise variance $\sigma_v^2$ is chosen such that a \gls{SNR} 
\begin{equation}
  \SNR = \frac{\sqrt{\inv{N}\sum_n y_{\Delta0}^2(n) }}%
              {\sqrt{\inv{N}\sum_n v^2(n) }}
       = \frac{\sigma_{y_{\Delta0}}}{\sigma_v}
       \approx \frac{\norm[2]{\Delta} \sigma_{u_{\Delta}}}{\sigma_v}
  \label{eq:SNR}
\end{equation}
of $10$ is obtained.

@@

We simulate the system \eqref{eq:O2sysDT} for $16\,100$ samples and discard the first $100$ samples to randomize the initial conditions.
To mimic the effect of progressively measuring longer data records, the first remaining $N$ samples are used to estimate $\infnorm{\Delta}$ as in \secref{sec:LPMHinf}.
The values of $N \in \ClosedInterval{35}{16 000}$ are chosen such that $\expectedValue{\gamma_{\mathrm{FRF}}(N)}$ is a non-decreasing function of $N$.
These simulations are then repeated in a Monte Carlo simulation of $n_{\mathrm{MC}}=200$ runs where each run has a different realization of  $u_{\Delta}(n)$ and $v(n)$.
The resulting sample average and standard deviation $\sigma$ are shown in \figref{fig:MC-run-length}.

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.75\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \input{\thisDir/figs/MC-runlength.tikz}
  %\input{figures/runLength.tikz}
  \caption{The $\Hinf{}$ norm~\legref{leg:runlength-true} is underestimated significantly for short experiments ($N<20\tau$) by $\expectedValue{\gamma_{\mathrm{FRF}}}$~\legref{leg:runlength-frf} and the \gls{LPM}~\legref{leg:runlength-LPM}.
  However, the proposed $\widehat{\norm[]{\Delta_{\mathrm{LRM}}}}_{\infty}$~\legref{leg:runlength-LRM} is already reliable from much shorter experiments ($N \geq 4 \tau$). A single observation for $N=97$~\legref{leg:runlength-special-case} is shown in \figref{fig:exampleIntergrid}. The shaded areas show the $\pm \sigma$ intervals which are mostly due to the noise.
  }
\label{fig:MC-run-length}
\end{figure}

It can be seen in \figref{fig:MC-run-length} that $N \geq 20\tau$ is required to attain a frequency grid that is dense enough to obtain a reliable $\gamma_{\mathrm{FRF}}$, i.e. $\expectedValue{\gamma_{\mathrm{FRF}}} \approx \gamma$.
Starting from about $N=4 \text{ to } 5\tau$, the interpolated \gls{LRM} starts to yield good estimates for $\gamma_{\mathrm{IG}}$.
For long experiments $(N \to \infty)$, the true $\norm{\Delta}$ is recovered.
Practically, this implies that the interpolated \gls{LRM} beats the on-grid estimate using a data record that is only a fraction of the length.

@@

\begin{figure}
 \centering
    \setlength{\figurewidth}{0.75\columnwidth}
    \setlength{\figureheight}{0.68\figurewidth}
    \input{\thisDir/figs/avis-frf.tikz}
 \caption{Measured \gls{FRF} \legref{m2t:plantFRF} and estimated parametric model $\hat{P}$ \legref{m2t:plantModel} of the \glsfirst{AVIS}.}
 \label{fig:avis-frf}
 \label{fig:avis}
\end{figure}

>>>>

\TODO{
  \begin{itemize}
    \item fix labels for chapter
    \item add section from Benchmark paper
    \item fix some bib entries
    \item Hinf macro aanpassen voor PDF support (indien dit niet conflicteert met glossaries)
  \end{itemize}
}
\input{\thisDir/styles.inc.tex}

@@

To estimate $\gamma$, only $\Delta(\omega_{\star})$ is of interest, where
$ \omega_{\star} 
     \isdef 
       \arg
         \sup_{\omega} 
           \abs{\Delta(\omega)}
  \text{.}
  \label{eq:omega-star}
$
However, in a finite measurement time, non-parametric approaches only give access to a finite and discrete grid of frequencies $\omega_k \in \DFTGrid$.
As such, \eqref{eq:HinfNormSISO} can be rewritten as
\begin{equation}
  \gamma \isdef \sup_{\omega} \abs{\Delta(\omega)}
         = \sup 
             \set{
                \ezbrace{\sup_{\omega \in \DFTGrid  } \abs{\Delta(\omega)}}{\gamma_{\mathrm{FRF}}}
                \;,\;
                \ezbrace{
                \sup_{\omega \in \InterGrid} \abs{\Delta(\omega)}}{\gamma_{\mathrm{IG}}}}
  \label{eq:gamma-split-grid}
\end{equation}
where $\Omega \isdef \OpenInterval{0}{\frac{2\pi}{\Ts}}$ is the measured frequency range for a sampling time $\Ts$.
Two cases can be distinguished:
\begin{enumerate}
  \item $\omega_{\star} \in \DFTGrid$ such that $\abs{\Delta(\omega_{\star})} = \gamma =\gamma_{\mathrm{FRF}}$ is observed directly on the discrete frequency grid,
  \item $\omega_{\star} \not\in \DFTGrid$, i.e. the extremal value $\abs{\Delta(\omega_{\star})}  = \gamma = \gamma_{\mathrm{IG}}$ is determined by the inter-grid behavior.
\end{enumerate}
The use of $\gamma_{\mathrm{FRF}}$ is standard in \gls{FRF}-based approaches, e.g.~\citep{vandeWal2002}.
Its expected value over different realizations of the input signal $\expectedValue{\gamma_{\mathrm{FRF}}} \leq \gamma$ due to the limited frequency resolution.
Such an underestimate may lead to overly optimistic controller designs.

@@

\section{\Hinf{} Norm Estimation Procedure}
\label{sec:LPMHinf}
To estimate $\gamma_{\mathrm{IG}}$, we propose to identify local parametric models in the frequency domain.
The frequency axis is partitioned into continuous segments $\Omega_k \isdef \OpenInterval{ \omega_{k-1} }{ \omega_{k+1} }$ in view of the local validity of the local models.
In particular, the inter-grid aspect of the $\Hinf$ norm is estimated as
\begin{equation}
  \gamma_{\mathrm{IG}} = 
    \sup \set{ \left. 
                \sup_{\omega \in \Omega_k} \abs{\LocalModel{\Delta}_k(\omega)}
                \right|
                k = 1,\ldots,N-1
                }
   \label{eq:gamma-inter-grid}
\end{equation}
where the $\LocalModel{\Delta}_{k}$ is the local model around the ($k$th) frequency $\omega_k$.
An example of such local models is shown in \figref{fig:interpol-quantities}.
Two alternative methods allow evaluation of~\eqref{eq:gamma-inter-grid} in practice.

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.8\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \input{\thisDir/figs/interpolation-show.tikz}
  \caption{Illustration of the local models $\LocalModel{\Delta}_k$ around each frequency bin $\omega_k$ for $k\in\set{1,\ldots,6}$.
  The bottom part shows the equivalence of the frequency segments $\Omega_k$ and the intervals $\set{\omega \given{k_L(\omega) = k \vee k_R(\omega) = k}}$.}
  \TODO{top subplot has a line connecting the last to the first point...}
\label{fig:interpol-quantities}
\end{figure}

@@

\section{Simulation}
\label{sec:simulation}
To allow the generalization of the results of the simulations, we first introduce a way to decompose complex systems into simpler systems.
The behavior of these simpler systems then allows to predict the behavior for these more complex systems.
Any real-rational system $\Delta$ can be decomposed in a parallel structure of lower-order sub-systems by means of a partial fraction expansion as in \eqref{eq:excitation:PFE} and using the same logic as in \chapref{sec:excitation}.
As a result, systems can often be well-represented, in a small frequency range, by a sum of a few low-order sub-systems (often first or second order for e.g. mechanical systems)~\citep[Section 2.2.2]{Gawronski2004}.
To estimate $\norm{\Delta}$ mainly the resonant sub-systems are of interest.
Assuming such resonances are sufficiently separated in the frequency,~\citep{Schoukens2013LPMerror,Geerardyn2013TIM,Gawronski2004} show that results for a single resonance are applicable to more complex systems.
This is done by considering only the dominant resonance -- \emph{viz.} the one that is the most relevant for the $\Hinf{}$ controller design.
By normalizing the conclusions for the time constant $\timeconst$ (or the $3\unit{dB}$ bandwidth), the results are generally valid for different values of the damping $\damping$.
Specifically, the system's $\timeconst \isdef (\damping \wn)^{-1}$ dictates the required measurement time~\citep{Schoukens2013LPMerror}, where $\damping$ is the relative damping and $\wn$ the undamped natural frequency.

\subsection{A concrete simulation}
Consider a discrete time system $\Delta$ (see \figref{fig:exampleIntergrid}) as unmodeled dynamics in an output-error setting:
\begin{equation}
  \Delta(z) 
    =  \frac{0.45373 z + 0.44752}
            {z^2 - 1.0595 z + 0.96079}
  \label{eq:O2sysDT}
  \text{,}
\end{equation}
which has $\timeconst = 50 \unit{samples}$ (i.e. $\damping=0.02$ and $\wn = 1 \unit{rad/s}$).
The input signal $u_{\Delta}$ is white Gaussian noise with unit variance and the disturbing noise variance $\sigma_v^2$ is chosen such that a \gls{SNR} 
\begin{equation}
  \SNR = \frac{\sqrt{\inv{N}\sum_n y_{\Delta0}^2(n) }}%
              {\sqrt{\inv{N}\sum_n v^2(n) }}
       = \frac{\sigma_{y_{\Delta0}}}{\sigma_v}
       \approx \frac{\norm[2]{\Delta} \sigma_{u_{\Delta}}}{\sigma_v}
  \label{eq:SNR}
\end{equation}
of $10$ is obtained.

@@

We simulate the system \eqref{eq:O2sysDT} for $16\,100$ samples and discard the first $100$ samples to randomize the initial conditions.
To mimic the effect of progressively measuring longer data records, the first remaining $N$ samples are used to estimate $\infnorm{\Delta}$ as in \secref{sec:LPMHinf}.
The values of $N \in \ClosedInterval{35,16\,000}$ are chosen such that $\expectedValue{\gamma_{\mathrm{FRF}}(N)}$ is a non-decreasing function of $N$.
These simulations are then repeated in a Monte Carlo simulation of $n_{\mathrm{MC}}=200$ runs where each run has a different realization of  $u_{\Delta}(n)$ and $v(n)$.
The resulting sample average and standard deviation $\sigma$ are shown in \figref{fig:MC-run-length}.

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.75\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \input{\thisDir/figs/MC-runlength.tikz}
  %\input{figures/runLength.tikz}
  \caption{The $\Hinf{}$ norm~\legref{leg:runlength-true} is underestimated significantly for short experiments ($N<20\timeconst$) by $\expectedValue{\gamma_{\mathrm{FRF}}}$~\legref{leg:runlength-frf} and the \gls{LPM}~\legref{leg:runlength-LPM}.
  However, the proposed $\widehat{\norm[]{\Delta_{\mathrm{LRM}}}}_{\infty}$~\legref{leg:runlength-LRM} is already reliable from much shorter experiments ($N \geq 4 \timeconst$). A single observation for $N=97$~\legref{leg:runlength-special-case} is shown in \figref{fig:exampleIntergrid}. The shaded areas show the $\pm \sigma$ intervals which are mostly due to the noise.
  }
\label{fig:MC-run-length}
\end{figure}

It can be seen in \figref{fig:MC-run-length} that $N \geq 20\timeconst$ is required to attain a frequency grid that is dense enough to obtain a reliable $\gamma_{\mathrm{FRF}}$, i.e. $\expectedValue{\gamma_{\mathrm{FRF}}} \approx \gamma$.
Starting from about $N=4 \text{ to } 5\timeconst$, the interpolated \gls{LRM} starts to yield good estimates for $\gamma_{\mathrm{IG}}$.
For long experiments $(N \to \infty)$, the true $\norm{\Delta}$ is recovered.
Practically, this implies that the interpolated \gls{LRM} beats the on-grid estimate using a data record that is only a fraction of the length.

@@

\begin{figure}
  \centering
  \setlength\figurewidth{0.75\columnwidth}
  \includegraphics[width=\figurewidth]{\thisDir/figs/avis.jpg}
  \caption{\Glsentryfirst{AVIS}.}
  \label{fig:avis}
\end{figure}

\begin{figure}
 \centering
    \setlength{\figurewidth}{0.75\columnwidth}
    \setlength{\figureheight}{0.68\figurewidth}
    \input{\thisDir/figs/avis-frf.tikz}
 \caption{Measured \gls{FRF} \legref{m2t:plantFRF} and estimated parametric model $\hat{P}$ \legref{m2t:plantModel} of the \glsfirst{AVIS}.}
 \label{fig:avis-frf}
\end{figure}
